---
title: "sealevel_VH"
author: "Kenneth Wong"
date: "10/28/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

library(raster)
library(rayshader)
library(rayrender)
library(rgdal)
library(magick)

library(gifski)
```

```{r}
VH_5m <- raster::raster("data/data-raw/VictoriaHarbour_5m_rect.tif")

VH_10m <- aggregate(VH_5m, 2)

VH_10m_matrix <- raster_to_matrix(VH_10m)
```

Original 10m DTM

```{r}

demold_10m_VH <- raster::raster("data/data-raw/demold_10m_VH.tif")
VH_10m_matrix <- raster_to_matrix(demold_10m_VH)

```

```{r}
VH_10m_matrix %>%
  sphere_shade(texture = "imhof1") %>%
  plot_map()
```


Lands Department 5m DTM, aggregated to 10m resolution in ArcGIS

```{r}

Lands_10m_VH <- raster::raster("data/data-raw/Lands_10m_VH.tif")
VH_10m_matrix <- raster_to_matrix(Lands_10m_VH)

```

Smoothed Lands DTM

Definition of 0m in DTM: 0 meter above HK Principal Datum
`MEAN_SEA_LEVEL`: 1.30 mPD

Heights are shown in metres above the Hong Kong Principal Datum (HKPD). For general reference, the Mean Sea Level is about 1.3 metres above the HKPD.
(https://www.landsd.gov.hk/mapping/en/download/data/iB50000_Data_Dictionary-GMLv.1.0.pdf)


```{r}

Lands_10m_smooth_VH <- raster::raster("data/data-mid/Lands_10m_VH_gaus_sdp5_r2.tif")

VH_10m_matrix <- raster_to_matrix(Lands_10m_smooth_VH)

```



Combined DTM (Lands 5m + spot height)
```{r}

edited_5m_VH <- raster::raster("data/data-raw/edited_5m_VH.tif")

edited_10m_VH <- raster::aggregate(edited_5m_VH, 2)

VH_10m_matrix <- raster_to_matrix(edited_10m_VH)

```

```{r}
VH_10m_matrix %>%
  sphere_shade(texture = "imhof1") %>%
  plot_map()
```

---

## Compute shadow


```{r}
VH_10m_shadow_z1 <- ray_shade(VH_10m_matrix, zscale = 1, lambert = TRUE)
VH_10m_amb_z1 <- ambient_shade(VH_10m_matrix, zscale = 1)




par(mfrow = c(1, 2))

plot_map(VH_10m_shadow_z1)
plot_map(VH_10m_amb_z1)
```


---

## static view

```{r}

ZSCALE <- 7.5

ZSCALE_SHADOW <- 1.5

VH_10m_shadow <- ray_shade(VH_10m_matrix, zscale = ZSCALE_SHADOW, lambert = TRUE)
VH_10m_amb <- ambient_shade(VH_10m_matrix, zscale = ZSCALE_SHADOW)
```

View shadows
```{r}
par(mfrow = c(1, 2))

plot_map(VH_10m_shadow)
plot_map(VH_10m_amb)
```



Shadow opacity tests

```{r}
par(mfrow = c(1, 2))

VH_10m_matrix %>%
  sphere_shade(zscale = 3, texture = "imhof1") %>%
  add_water(
    detect_water(VH_10m_matrix, min_area = length(VH_10m_matrix)/100),
    color = "turquoise4"
    ) %>%
  add_shadow(VH_10m_shadow, .5) %>%
  add_shadow(VH_10m_amb, .25) %>%
  plot_map()

VH_10m_matrix %>%
  sphere_shade(zscale = 3, texture = "imhof1") %>%
  add_water(
    detect_water(VH_10m_matrix, min_area = length(VH_10m_matrix)/100),
    color = "turquoise4"
    ) %>%
  add_shadow(VH_10m_shadow, .4) %>%
  add_shadow(VH_10m_amb, .2) %>%
  plot_map()

```



```{r}
WATER_DEPTH = 4.9

VH_10m_matrix %>%
  sphere_shade(texture = "imhof1", sunangle = 315) %>%
  add_shadow(VH_10m_shadow, .75) %>%
  add_shadow(VH_10m_amb, .25) %>%
  plot_3d(VH_10m_matrix, zscale = ZSCALE, fov = 0, theta = -45, phi = 35,
            windowsize = c(1920, 1200), zoom = 0.6,
            solid = TRUE, soliddepth = -50, solidcolor = "grey20")

render_snapshot()


render_water(VH_10m_matrix, zscale = ZSCALE,
             waterdepth = WATER_DEPTH, wateralpha = 0.5, watercolor = "turquoise4",
             waterlinecolor = "#ffffff", waterlinealpha = 0.5, linewidth = 4)

# "VH_Axo_Zoomtest.png"

render_snapshot(
                title_text = paste(WATER_DEPTH, "m"),
                title_position = "northeast",
                title_font = "Helvetica",
                title_size = 72,
                clear = TRUE)

```

### Top

```{r}
WATER_DEPTH = 0.1

VH_10m_matrix %>%
  sphere_shade(texture = "imhof1", sunangle = 315) %>%
  add_shadow(VH_10m_shadow, .8) %>%
  add_shadow(VH_10m_amb, .25) %>%
  plot_3d(VH_10m_matrix, zscale = ZSCALE, fov = 0, theta = 0, phi = 90, 
            windowsize = c(2560, 1600), zoom = 0.65)
            # solid = TRUE, soliddepth = -50, solidcolor = "grey20")

render_snapshot()
# render_snapshot("VH_noWwater.png", clear = TRUE)

# No need to render waterline for top view
render_water(VH_10m_matrix, zscale = ZSCALE,
             waterdepth = WATER_DEPTH, wateralpha = 0.5, watercolor = "turquoise4",
             waterlinecolor = NULL)

# "5p1m_Lands10m.png",

render_snapshot(
                title_text = paste(WATER_DEPTH, "m"),
                title_position = "northeast",
                title_font = "Helvetica",
                title_size = 90,
                clear = TRUE)

```

---

## Notes on sea level

In edited hybrid model, base sea level = 2
In Lands DTM, base sea level = 0

---

## Animation

### Animation params

From https://wcmbishop.github.io/rayshader-demo/

#### For edited DTM

```{r}
FRAMES_NUM <- 120

ZSCALE <- 7.5

RISE_METERS <- 5

# In edited DTM, the base sea level about 2m (assume it equal to 0mPD)
MEAN_SEA_LEVEL <- 2

# Plus epsilon (0.01-0.1) since waterdepth at 0 will "interlace" with the sea surface
EPSILON <- 0.1

# Smooth sine/cosine raise
# Plus 0.1 since waterdepth at 0 will "interlace" with the sea surface
# length.out = desired length of the sequence
WATERDEPTHS_SMOOTH <- RISE_METERS/2 - RISE_METERS/2 * cos(seq(0, 2*pi, length.out = FRAMES_NUM)) + MEAN_SEA_LEVEL + EPSILON

# WATERDEPTHS_SMOOTH_2 <- 10 * sin(seq(0, pi, length.out = FRAMES_NUM)) + 2.1

# Uniform raising speed
WATERDEPTHS_RISE <- seq(0, 1, length.out = FRAMES_NUM/2) * RISE_METERS + MEAN_SEA_LEVEL + EPSILON

# rev() to reverse sequence order (i.e. decreasing water level)
WATERDEPTHS_UNISPEED <- c(WATERDEPTHS_RISE, rev(WATERDEPTHS_RISE))


# From 315 (-45) to 45
THETAS <- 45 * -cos(seq(0, 2*pi, length.out = FRAMES_NUM))

```


#### For Lands 10m DTM (both smoothed/non-smoothed)

Changes in `MEAN_SEA_LEVEL` and `EPSILON`

Notes on definition of sea level
http://www.geodetic.gov.hk/smo/gsi/data/pdf/explanatorynotes_c.pdf


```{r}
FRAMES_NUM <- 120

ZSCALE <- 7.5

# Lands DTM use HKPD as 0
# Mean Sea Level defined by SMO is 1.30 mPD
MEAN_SEA_LEVEL <- 1.30

# Plus epsilon since waterdepth at 0m will "interlace" with the sea surface
EPSILON <- 0.01

# meters rise from current sea level
RISE_METERS <- 5

# Smooth sine/cosine raise
# length.out = desired length of the sequence
WATERDEPTHS_SMOOTH <- RISE_METERS/2 - RISE_METERS/2 * cos(seq(0, 2*pi, length.out = FRAMES_NUM)) + MEAN_SEA_LEVEL + EPSILON

# WATERDEPTHS_SMOOTH_2 <- 10 * sin(seq(0, pi, length.out = FRAMES_NUM)) + 2.1

# Uniform raising speed
WATERDEPTHS_RISE <- (seq(0, 1, length.out = FRAMES_NUM/2) * RISE_METERS) + MEAN_SEA_LEVEL + EPSILON

# rev() to reverse sequence order (i.e. decreasing water level)
WATERDEPTHS_UNISPEED <- c(WATERDEPTHS_RISE, rev(WATERDEPTHS_RISE))


# From 315 (-45) to 45
THETAS <- 45 * -cos(seq(0, 2*pi, length.out = FRAMES_NUM))

```

### Top view

If the sea level lowering part is just an identical mirror image to sea level rising, just need to render the first half, then play in reverse order when combining images

```{r, echo = FALSE}

# generate .png frame images
img_frames <- paste0("VH_Top_uni_Lands10m_smooth/VHTop_", seq_len(FRAMES_NUM), ".png")

for (i in seq_len(FRAMES_NUM/2)) {
  message(paste(" - Generating image", i, "of", FRAMES_NUM/2))
  
  message(paste("     ", WATERDEPTHS_UNISPEED[i] - (MEAN_SEA_LEVEL + EPSILON), "m rise in sea level"))
  
  start_time <- Sys.time()

  VH_10m_matrix %>%
    sphere_shade(texture = "imhof1", sunangle = 315) %>%
    add_shadow(VH_10m_shadow, .75) %>%
    add_shadow(VH_10m_amb, .25) %>%
    plot_3d(VH_10m_matrix, zscale = ZSCALE, fov = 0, theta = 0, phi = 90, 
            windowsize = c(2560, 1600), zoom = 0.65)
  
  render_snapshot()
  
  # Add water
  render_water(VH_10m_matrix, zscale = ZSCALE,
               waterdepth = WATERDEPTHS_UNISPEED[i], wateralpha = 0.5, watercolor = "turquoise4",
               # No need to render waterline for top view
               waterlinecolor = NULL)

  
  # Render final view
  render_snapshot(img_frames[i],
                  # In DTM, the base sea level about 2.1m (assume it equal to 0mPD)
                  # Fix to always have 2 decimals
                  # https://stackoverflow.com/questions/3443687/formatting-decimal-places-in-r
                  title_text = paste("+", format(round(WATERDEPTHS_UNISPEED[i] - (MEAN_SEA_LEVEL + EPSILON), 2), nsmall = 2), "m"),
                  title_position = "northeast",
                  title_font = "Helvetica",
                  title_size = 90,
                  clear = TRUE)
  
  rgl::clear3d()
  
  end_time <- Sys.time()
  
  print(end_time - start_time)
}


```

```{r}
# generate .png frame images


img_frames_rise <- paste0("VH_Top_uni_Lands10m_smooth/VHTop_", seq_len(FRAMES_NUM/2), ".png")

img_frame_sequence <- c(
  # rising
  img_frames_rise,
  # stop at top for 0.5 sec
  rep(paste0("VH_Top_uni_Lands10m_smooth/VHTop_", FRAMES_NUM/2, ".png"), 12),
  # dropping
  rev(img_frames_rise),
  # stop at bottom for 0.5 sec
  rep(paste0("VH_Top_uni_Lands10m_smooth/VHTop_", 1, ".png"), 12))

# write video
magick::image_write_video(magick::image_read(img_frame_sequence), 
                          path = "VH_sealevel_Top_uni_Lands10m_smooth.mp4", 
                          framerate = 12)
```



### Axo view

```{r, echo = FALSE}

# generate .png frame images
img_frames <- paste0("VH_Axo_uni/VHAxo_", seq_len(FRAMES_NUM), ".png")

for (i in seq_len(FRAMES_NUM/2)) {
  message(paste(" - Generating image", i, "of", FRAMES_NUM))
          
  message(paste(" - Sea level of", format(round(WATERDEPTHS_UNISPEED[i] - (MEAN_SEA_LEVEL + EPSILON), 2), nsmall = 2), "m"))
  
  start_time <- Sys.time()

  VH_10m_matrix %>%
    sphere_shade(texture = "imhof1", sunangle = 315) %>%
    add_shadow(VH_10m_shadow, .75) %>%
    add_shadow(VH_10m_amb, .25) %>%
    plot_3d(VH_10m_matrix, zscale = ZSCALE, fov = 0, theta = -45, phi = 35,
            windowsize = c(1920, 1200), zoom = 0.6,
            solid = TRUE, soliddepth = -50, solidcolor = "grey20")

  render_snapshot()
  
  # Add water
  render_water(VH_10m_matrix, zscale = ZSCALE,
               waterdepth = WATERDEPTHS_UNISPEED[i], wateralpha = 0.5, watercolor = "turquoise4",
               waterlinecolor = "#ffffff", waterlinealpha = 0.5, linewidth = 4)
  
  # Render final view
  render_snapshot(img_frames[i],
                  
                  # Fix to always have 2 decimals
                  # https://stackoverflow.com/questions/3443687/formatting-decimal-places-in-r
                  title_text = paste(format(round(WATERDEPTHS_UNISPEED[i] - (MEAN_SEA_LEVEL + EPSILON), 2), nsmall = 2), "m"),
                  title_position = "northeast",
                  title_font = "Helvetica",
                  title_size = 72,
                  clear = TRUE)
  
  rgl::clear3d()
  
  end_time <- Sys.time()
  
  print(end_time - start_time)
}


# write video
magick::image_write_video(magick::image_read(img_frames), 
                          path = "VH_sealevel_Axo_uni_editedDTM.mp4", 
                          framerate = 12)
```

```{r}
# generate .png frame images
img_frames_rise <- paste0("VH_Axo_uni_editedDTM/VHAxo_", seq_len(FRAMES_NUM/2), ".png")

img_frame_sequence <- c(
  # rising
  img_frames_rise,
  # stop at top for 0.5 sec
  rep(paste0("VH_Axo_uni_editedDTM/VHAxo_", FRAMES_NUM/2, ".png"), 12),
  # dropping
  rev(img_frames_rise),
  # stop at bottom for 0.5 sec
  rep(paste0("VH_Axo_uni_editedDTM/VHAxo_", 1, ".png"), 12))

# write video
magick::image_write_video(magick::image_read(img_frame_sequence), 
                          path = "VH_sealevel_Axo_uni_editedDTM.mp4", 
                          framerate = 12)
```

### Axo view, changing view angle

```{r, echo = FALSE}

# generate .png frame images
img_frames <- paste0("VH_Axo_uni_theta/VHAxoTheta_", seq_len(FRAMES_NUM), ".png")

for (i in seq_len(FRAMES_NUM)) {
  message(paste(" - Generating image", i, "of", FRAMES_NUM))
  
  start_time <- Sys.time()

  VH_10m_matrix %>%
    sphere_shade(texture = "imhof1", sunangle = 315) %>%
    add_shadow(VH_10m_shadow, .75) %>%
    add_shadow(VH_10m_amb, .25) %>%
    plot_3d(VH_10m_matrix, zscale = ZSCALE, fov = 0, theta = THETAS[i], phi = 35,
            windowsize = c(1920, 1200), zoom = 0.6,
            solid = TRUE, soliddepth = -50, solidcolor = "grey20")

  
  render_snapshot()
  
  # Add water
  render_water(VH_10m_matrix, zscale = ZSCALE,
               waterdepth = WATERDEPTHS_UNISPEED[i], wateralpha = 0.5, watercolor = "turquoise4",
               waterlinecolor = "#ffffff", waterlinealpha = 0.5, linewidth = 4)
  
  # Render final view
  render_snapshot(img_frames[i],
                  # In DTM, the base sea level about 2.1m (assume it equal to 0mPD)
                  # Fix to always have 2 decimals
                  # https://stackoverflow.com/questions/3443687/formatting-decimal-places-in-r
                  title_text = paste(format(round(WATERDEPTHS_SMOOTH[i] - 2.1, 2), nsmall = 2), "m"),
                  title_position = "northeast",
                  title_font = "Helvetica",
                  title_size = 72,
                  clear = TRUE)
  
  rgl::clear3d()
  
  end_time <- Sys.time()
  
  print(end_time - start_time)
}


# write video
magick::image_write_video(magick::image_read(img_frames), 
                          path = "VH_sealevel_Axo_uni_phi_f30.mp4", 
                          framerate = 30)
```




---

## Area affected


```{r, eval = FALSE}
# Directly plot histogram with raster

Lands_10m_above_sea_level <- Lands_10m_smooth_VH

Lands_10m_above_sea_level[Lands_10m_above_sea_level <= 1.30] <- NA

raster::hist(Lands_10m_above_sea_level, maxpixels = 1e5, breaks = seq(-25, 600, 5))
```

```{r}
# convert to df
# better option to deal with ggplot

height_df <- raster::as.data.frame(Lands_10m_smooth_VH)

colnames(height_df) <- c("height")
```


```{r}
# Lands DTM use HKPD as 0
# Mean Sea Level defined by SMO is 1.30 mPD
MEAN_SEA_LEVEL <- 1.30

height_df_10m_above <- height_df %>%
  filter(height > MEAN_SEA_LEVEL & height <= (MEAN_SEA_LEVEL + 10)) %>%
  mutate(height_above_sea = height - MEAN_SEA_LEVEL)
```


```{r}
# Histogram
ggplot(height_df_10m_above, aes(x = height_above_sea)) +
  geom_histogram(binwidth = .05)

# Cumulative frequency
ggplot(height_df_10m_above, aes(x = height_above_sea)) +
  stat_ecdf() +
  theme_bw()
```

```{r}
# Cumulative counts
# https://stackoverflow.com/questions/18379933/plotting-cumulative-counts-in-ggplot2

ggplot(height_df_10m_above, aes(x = height_above_sea)) +
  stat_bin(aes(y = cumsum(..count..) * 100 / 1e4), geom = "step", binwidth = .25, color = "#00868b") +
  theme_bw() +
  theme(
    panel.spacing = unit(.1, "lines"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.text = element_text(size = 8),
    axis.title.x = element_text(size = 11),
    axis.title.y = element_text(size = 11),
    
    # https://stackoverflow.com/questions/41105759/aligning-title-subtitle-and-caption-for-horizontal-ggplot-barchart
    plot.caption = element_text(hjust = 0),
    plot.caption.position = "plot"
  ) +
  xlim(c(0, 10)) +
  labs(
    title = "Total flooded area when sea level rises from 0 to 10m",
    x = "Rise from mean sea level (m)",
    y = "Total flooded area (ha)",
    caption = "Mean sea level is defined as 1.30m above Hong Kong Principal Datum (HKPD)."
  )

```


```{r}

# size of common place for comparison, in ha

HIGH_ISLAND_RESORVOIR <- 541
VICTORIA_PARK <- 19

# Consider add vertical line the cumulative plot
# https://stackoverflow.com/questions/12876501/r-ggplot2-labelling-a-horizontal-line-on-the-y-axis-with-a-numeric-value
# + geom_hline(aes(yintercept = h))
```



